---
title: "Erson Analysis Vignette"
author: "Sebastian Reyes"
date: "2025-05-16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
```

## 0. Knitting this Document

```{r, eval = FALSE}
# this chunk knits the document
# if all necessary objects are not loaded into the environment, clear it and run the chunk below prior to knitting
rmarkdown::render("Erson_Analysis_Vignette.Rmd", envir = globalenv())

```

```{r, include = FALSE, eval = FALSE}
# if all necessary objects are not loaded into environment, clear it and then run this chunk prior to knitting
#this code is also integrated below into the function examples themselves

# Functions
#To create an inclusive barplot based on cell identities 
CreateSeuratObject <- function(fileDirectory, objectName, projectName) {
  rawCounts <- Read10X_h5(fileDirectory)
  seurat_obj <- Seurat::CreateSeuratObject(counts = rawCounts, 
                                           project = projectName, 
                                           min.cells = 3, 
                                           min.features = 200)
  assign(objectName, seurat_obj, envir = .GlobalEnv)
  return(objectName)
}

AddMitoRna <- function(seuratObject){
  seuratObject[["percent.mt"]] <- Seurat::PercentageFeatureSet(seuratObject, pattern = "^MT-")
  return(seuratObject)
}

MakeVlnPlot <- function(seuratObject) {
  print(VlnPlot(seuratObject, 
                features = c("nFeature_RNA", 
                             "nCount_RNA", 
                             "percent.mt"), 
                ncol = 3))
  return(seuratObject)
}

QCSeurat <- function(seuratObject, minFeature, maxFeature, percentMT) {
  seuratObject <- subset(seuratObject, subset = nFeature_RNA > minFeature & nFeature_RNA < maxFeature & percent.mt < percentMT)
  return(seuratObject)
}

NormalizeAndVariable <- function(seuratObject) {
  seuratObject <- NormalizeData(seuratObject)
  seuratObject <- FindVariableFeatures(seuratObject, selection.method = "vst", nfeatures = 2000)
  top10 <- head(VariableFeatures(seuratObject), 5)
  plot1 <- VariableFeaturePlot(seuratObject)
  plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
  print(plot2)
  return(seuratObject)
}

Scale <- function(seuratObject) {
  seuratObject <- ScaleData(seuratObject, features = rownames(seuratObject))
  seuratObject <- ScaleData(seuratObject, vars.to.regress = "percent.mt",)
  return(seuratObject)
}

PcaElbow <- function(seuratObject) {
  seuratObject <- RunPCA(seuratObject, features = VariableFeatures(seuratObject))
  print(ElbowPlot(object = seuratObject))
  return(seuratObject)
}

MakeHeatmap <- function(seuratObject, PcNumberVector){
  DimHeatmap(seuratObject, dims = PcNumberVector, balanced = TRUE)
  return(seuratObject)
}

MakeUMAP <- function(seuratObject1, PCs) {
  seuratObject2 <- RunUMAP(seuratObject1, dims = PCs)
  print(DimPlot(seuratObject2, reduction = "umap"))
  return(seuratObject2)
}

FindPosMarkers <- function(seuratObject){
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  posMarkers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC >1)
  return(seuratObject)
}

Top5Markers <- function(seuratObject) {
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  separatedMarkers <- posMarkers %>%
    dplyr::filter(avg_log2FC >1) %>%
    group_by(cluster) %>%
    slice_head(n = 5)
  print(separatedMarkers, n = Inf)
}

SingleHPCA <- function(seuratObject){
  HpcaLabels <- SingleR(test = as.SingleCellExperiment(seuratObject), 
                        ref =    HumanPrimaryCellAtlasData(), 
                        labels = HumanPrimaryCellAtlasData()$label.main)
  seuratObject1 <- AddMetaData(seuratObject, 
                               metadata = HpcaLabels@listData[["labels"]], 
                               col.name = "HPCACellLabels")
  return(seuratObject1)
}

SingleNeuro <- function(seuratObject) {
  # Function to run SingleR
  NeuroClusterLabels <- SingleR(test = as.SingleCellExperiment(seuratObject),
                                ref = NeuroExpressionFiltered,
                                labels = NeuroIDs)
  
  #To add SingleR output to Seurat Object
  seuratObject1 <- AddMetaData(seuratObject, metadata = NeuroClusterLabels@listData[["labels"]], col.name = "NeuroCellLabels")
  
  #To group everything by Neuro Cell Type
  Idents(seuratObject1) <- "NeuroCellLabels"
  return(seuratObject1)
}

SingleImmune <- function(seuratObject) {
  
SeuratObjectExpressionData <- as.matrix(GetAssayData(seuratObject, assay = "RNA", layer = "data"))

results <- SCINA(SeuratObjectExpressionData, ImmuneRef, rm_overlap = FALSE)
seuratObject1 <- AddMetaData(seuratObject, metadata = results$cell_labels, col.name = "ImmuneCellLabels")
Idents(seuratObject1) <- "ImmuneCellLabels"
return(seuratObject1)
}

SingleUnbiasedNeuro <- function(seuratObject){
  
  # Create a table of `seurat_clusters` vs `NeuroCellLabels`
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$NeuroCellLabels)
  
  # Identify the most abundant label for each cluster
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] # Returns the label with the highest count
  })
  
  # Relabel `seurat_clusters` with the most abundant `NeuroCellLabels`
  seuratObject@meta.data$UnbiasedNeuroCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  return(seuratObject)
}

SingleUnbiasedImmune <- function(seuratObject){
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$ImmuneCellLabels)
  
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] 
  })
  
  seuratObject@meta.data$UnbiasedImmuneCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  
  return(seuratObject)
}

BarPlotIdPercent <- function(seuratObject, Identity){
  
  TotalCells <- dim(seuratObject@meta.data)[1]
  
  cellIdentities <- seuratObject@meta.data[[Identity]]
  
  cellCounts <- table(cellIdentities)
  
  cellCountsDF <- as.data.frame(cellCounts)
  
  colnames(cellCountsDF) <- c("CellIdentity", "Count")
  
  ggplot(cellCountsDF, aes(x = CellIdentity, 
                           y = Count/TotalCells*100, 
                           fill = CellIdentity)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + 
    labs(title = "Distribution of Cell Types", x = "Cell Identity", y = "Percentage of Cells") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme(text = element_text("sans"))
}


# for MakeVlnPlot() example

CreateSeuratObject("DATA1.H5",
                   "ex_sample1", 
                   "ex_sample1")

ex_sample1 <- AddMitoRna(ex_sample1)

#for QCSeurat() example
ex_sample2 <- ex_sample1

#for NormalizeAndVariable() example
ex_sample3 <- ex_sample2
ex_sample3 <- QCSeurat(ex_sample3, 200, 3000, 25)

#for PcaElbow() example
ex_sample4 <- ex_sample3
ex_sample4 <- NormalizeAndVariable(ex_sample4)
ex_sample4 <- Scale(ex_sample4)

#for MakeHeatMap() example
ex_sample5 <- ex_sample4
ex_sample5 <- PcaElbow(ex_sample5)

#for FindCellClusters() example
ex_sample6 <- ex_sample5

#for MakeUMAP() example
ex_sample7 <- ex_sample6
ex_sample7 <- FindCellClusters(ex_sample7, 1:7, 0.5)

#for Top5Markers() example
ex_sample8 <- ex_sample7
ex_sample8 <- RunUMAP(ex_sample8, dims = 1:7)

#for SingleHPCA() example
ex_sample9 <- ex_sample8
ex_sample9 <- SingleHPCA(ex_sample9)

#for SingleNeuro() example
NeuroMeta <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_CellInfo.csv.gz")

NeuroExpression <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_5181_count.csv.gz")

columns_to_keep <- colnames(NeuroExpression) %in% NeuroMeta$X | colnames(NeuroExpression) == "X"

NeuroExpressionFiltered <- NeuroExpression[, columns_to_keep]

rm(NeuroExpression)

rownames(NeuroExpressionFiltered) <- NeuroExpressionFiltered[,1]
NeuroExpressionFiltered <- NeuroExpressionFiltered[,-1]

NeuroExpressionFiltered <- NeuroExpressionFiltered %>%
  mutate(across(everything(), as.numeric))

NeuroIDs <- as.character(NeuroMeta$cell_type)
names(NeuroIDs) <- colnames(NeuroExpressionFiltered)

rm(NeuroMeta)
NeuroExpressionFiltered <- as.matrix(NeuroExpressionFiltered)

ex_sample10 <- ex_sample9
ex_sample10 <- SingleNeuro(ex_sample10)

#for SingleImmune() example

ImmuneRef <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Immune Cells/E-CURD-122.marker_genesymbols_inferred_cell_type_SCINA.csv")

ImmuneRef_list <- lapply(1:39, function(i) {
  genes <- ImmuneRef[, i]
  genes <- genes[!is.na(genes)]
  return(genes)
})

names(ImmuneRef_list) <- colnames(ImmuneRef)
ImmuneRef <- ImmuneRef_list
rm(ImmuneRef_list)

ex_sample10 <- SingleImmune(ex_sample10)
ex_sample10imm <- subset(ex_sample10, NeuroCellLabels == "Imm")

CategoryCounts <- ex_sample10imm@meta.data %>%
  group_by(ImmuneCellLabels) %>%
  tally()

ValidCategories <- CategoryCounts %>% 
  filter(n >= 5) %>%
  pull(ImmuneCellLabels)

#for SingleUnbiasedNeuro() example
ex_sample10 <- SingleUnbiasedNeuro(ex_sample10)

#for SingleUnbiasedImmune() example
ex_sample10imm <- SingleUnbiasedImmune(ex_sample10imm)

```

## 1. Introduction

The below framework outlines a workflow I designed to streamline the steps outlined in this [Seurat vignette](https://satijalab.org/seurat/articles/pbmc3k_tutorial) for the purposes needed by the Erson Lab in the summer of 2024--when I made this. Really, the entire framework below is a reframing of this vignette to better suit the specific needs of the Erson Lab—from the perspective of my specific project, anyway.

To explain the "custom functions" created below: many of these functions are used repeatedly with different parameters during the exploratory analyses I conducted on the Erson Lab's samples. Making these custom functions quickened the workflow and allowed parameters to more easily be edited.

That said, reading the aforementioned [Seurat vignette](https://satijalab.org/seurat/articles/pbmc3k_tutorial) would provide the reader with the vast majority of the information contained in the below outline.

## 2. Installing Packages Needed

```{r, eval = FALSE}
install.packages("dplyr")
install.packages("Seurat")
install.packages("patchwork")
install.packages("tibble")
install.packages("SingleR")
install.packages("SCINA")
BiocManager::install("celldex")
install.packages("ggplot2")
```

## 3. Loading Packages Needed

```{r, eval = FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(tibble)
library(SingleR)
library(SCINA)
library(celldex)
library(ggplot2)
```

## 4. All Custom Functions Used

### CreateSeuratObject

#### **Explanation**:

-   More easily creates a Seurat object based on file directory
-   **Parameters**:
    -   fileDirectory: file path (must be h5 file for this specific function)

    -   objectName: self-explanatory

    -   projectName: self-explanatory

```{r, eval = FALSE}
CreateSeuratObject <- function(fileDirectory, objectName, projectName) {
  rawCounts <- Read10X_h5(fileDirectory)
  seurat_obj <- Seurat::CreateSeuratObject(counts = rawCounts, 
                                           project = projectName, 
                                           min.cells = 3, 
                                           min.features = 200)
  assign(objectName, seurat_obj, envir = .GlobalEnv)
  return(objectName)
}
```

### AddMitoRna

#### **Explanation**:

-   Appends mitochondrial RNA data to Seurat object on a cell-by-cell basis

```{r, eval = FALSE}
AddMitoRna <- function(seuratObject){
  seuratObject[["percent.mt"]] <- Seurat::PercentageFeatureSet(seuratObject, pattern = "^MT-")
  return(seuratObject)
}
```

### MakeVlnPlot

#### **Explanation**:

-   More easily creates violin plots for QC purposes

    -   nFeature_RNA = gives the number of unique features (genes) in a given cell, each point represents one cell

    -   nCount_RNA = gives the total number of RNA molecules in a cell, each point represents one cell

    -   percent.mt = percentage of mitochondrial RNA, each point represents one cell

```{r}
MakeVlnPlot <- function(seuratObject) {
  print(VlnPlot(seuratObject, 
                features = c("nFeature_RNA", 
                             "nCount_RNA", 
                             "percent.mt"), 
                ncol = 3))
  return(seuratObject)
}
```

#### **Example:**

```{r, eval = FALSE}
# Must be run previously:
CreateSeuratObject("DATA1.H5",
                   "ex_sample1", 
                   "ex_sample1")
ex_sample1 <- AddMitoRna(ex_sample1)
```

```{r}
MakeVlnPlot(ex_sample1)
```

### QCSeurat

#### **Explanation:**

-   Streamlines the quality control process, allows for low-quality cells to be eliminated based on feature count and mitochondrial RNA percentage

-   Parameters usually decided based on violin plots to minimize outliers/maximize viable retained data

-   **Parameters:**

    -   seuratObject = seuratObject to be trimmed

    -   minFeature = minimum number of genes a cell must contain to remain included in data

    -   maxFeature = maximum number of genes a cell can contain to remain included in data

    -   percentMT = maximum percentage of mitochondrial RNA which a cell can contain to remain included in data

```{r, eval = FALSE}
QCSeurat <- function(seuratObject, minFeature, maxFeature, percentMT) {
  seuratObject <- subset(seuratObject, subset = nFeature_RNA > minFeature & nFeature_RNA < maxFeature & percent.mt < percentMT)
  return(seuratObject)
}
```

#### Example:

```{r}
# Violin plot prior to QC
MakeVlnPlot(ex_sample2)

ex_sample2 <- QCSeurat(ex_sample2, 200, 3000, 25)

#Violin plot following QC
MakeVlnPlot(ex_sample2)
```

### NormalizeAndVariable

#### **Explanation**:

-   Normalizes data, finds genes within Seurat object that display variable expression patterns, and creates a plot highlighting the top 10 genes with the highest degree of variability in their expression.

```{r, eval = FALSE}
NormalizeAndVariable <- function(seuratObject) {
  seuratObject <- NormalizeData(seuratObject)
  seuratObject <- FindVariableFeatures(seuratObject, selection.method = "vst", nfeatures = 2000)
  top10 <- head(VariableFeatures(seuratObject), 5)
  plot1 <- VariableFeaturePlot(seuratObject)
  plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
  print(plot2)
  return(seuratObject)
}
```

#### Example:

```{r}
ex_sample3 <- NormalizeAndVariable(ex_sample3)
```

### Scale

#### **Explanation**:

-   Scales the Seurat object as per the workflow outlined in the Seurat vignette

```{r, eval = FALSE}
Scale <- function(seuratObject) {
  seuratObject <- ScaleData(seuratObject, features = rownames(seuratObject))
  seuratObject <- ScaleData(seuratObject, vars.to.regress = "percent.mt",)
  return(seuratObject)
}
```

### PcaElbow

#### **Explanation:**

-   Runs a principal component analysis and creates an elbow plot based on it

```{r, eval = FALSE}
PcaElbow <- function(seuratObject) {
  seuratObject <- RunPCA(seuratObject, features = VariableFeatures(seuratObject))
  print(ElbowPlot(object = seuratObject))
  return(seuratObject)
}
```

#### Example:

```{r, eval = FALSE}
# Must be run previously
ex_sample4 <- Scale(ex_sample4)
```

```{r, include = FALSE}
# DO NOT run the chunk below this manually. the first time it is run should be when the document is knitted
```

```{r}
ex_sample4 <- PcaElbow(ex_sample4)
```

### MakeHeatmap

#### Explanation

-   Creates one heatmap for every principal component included in `PcNumberVector`

    -   `PcNumberVector` should be a numeric vector

```{r, eval = FALSE}
MakeHeatmap <- function(seuratObject, PcNumberVector){
  DimHeatmap(seuratObject, dims = PcNumberVector, balanced = TRUE)
  return(seuratObject)
}
```

#### Example:

```{r}
# Note: PcaElbow must be run first, as it attaches PCA data to the Seurat object
MakeHeatmap(ex_sample5, 1:9)
```

### FindCellClusters

#### **Explanation:**

-   Finds clusters of cells based off principal component behavior

-   **Parameters**:

    -   seuratObject = self-explanatory

    -   PCs = a numeric vector containing the principal components to be used

    -   resolution = a value between 0 and 1; defines the granularity of the clusters

-   Cluster assignments may be found via the dataframe `seuratObject@meta.data`, column = `seurat_clusters`

```{r, eval = FALSE}
FindCellClusters <- function(seuratObject, PCs, resolution) {
  seuratObject2 <- FindNeighbors(seuratObject, dims = PCs)
  seuratObject3 <- FindClusters(seuratObject2, resolution = resolution)
  return(seuratObject3)
}
```

#### Example:

```{r}
# Below code finds 11 clusters
ex_sample6 <- FindCellClusters(ex_sample6, 1:7, 0.5)
```

### MakeUMAP

#### **Explanation:**

-   Runs UMAP + creates a UMAP plot based on the specified principal components

-   **Parameters:**

    -   seuratObject = self-explanatory

    -   PCs = a numeric vector containing the principal components to be used

```{r, eval = FALSE}
MakeUMAP <- function(seuratObject1, PCs) {
  seuratObject2 <- RunUMAP(seuratObject1, dims = PCs)
  print(DimPlot(seuratObject2, reduction = "umap"))
  return(seuratObject2)
}
```

#### Example:

```{r}
ex_sample7 <- MakeUMAP(ex_sample7, 1:7)
```

### FindPosMarkers

#### **Explanation:**

-   Finds markers that are upregulated within previously established clusters

```{r, eval = FALSE}
FindPosMarkers <- function(seuratObject){
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  posMarkers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC >1)
  return(seuratObject)
}
```

### Top5Markers

#### **Explanation:**

-   Prints out the top 5 most upregulated genetic markers on a cluster-by-cluster basis

```{r, eval = FALSE}
Top5Markers <- function(seuratObject) {
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  separatedMarkers <- posMarkers %>%
    dplyr::filter(avg_log2FC >1) %>%
    group_by(cluster) %>%
    slice_head(n = 5)
  print(separatedMarkers, n = Inf)
}
```

#### Example:

```{r}
# the actual output is 190 rows
head(Top5Markers(ex_sample8))
```

### SingleHPCA

#### **Explanation:**

-   Annotates each individual cell based on its expression profile's closest match within the Human Primary Cell Atlas reference dataset; based on the use of SingleR

    -   HPCA dataset is more generic than the below neuroendocrine or immune cell datasets, but is good for establishing an initial tissue composition overview

    -   Cell identities may be found via the dataframe `seuratObject@meta.data`, column = `HPCACellLabels`

```{r, eval = FALSE}
SingleHPCA <- function(seuratObject){
  HpcaLabels <- SingleR(test = as.SingleCellExperiment(seuratObject), 
                        ref =    HumanPrimaryCellAtlasData(), 
                        labels = HumanPrimaryCellAtlasData()$label.main)
  seuratObject1 <- AddMetaData(seuratObject, 
                               metadata = HpcaLabels@listData[["labels"]], 
                               col.name = "HPCACellLabels")
  return(seuratObject1)
}
```

#### Example:

```{r, eval = FALSE}
ex_sample9 <- SingleHPCA(ex_sample9)
```

```{r, include = FALSE}
# DO NOT run the chunk below this manually. the first time it is run should be when the document is knitted
```

```{r}
Idents(ex_sample9) <- ex_sample9@meta.data$HPCACellLabels
DimPlot(ex_sample9, reduction = "umap")
```

### SingleNeuro

-   **Explanation:** Annotates each individual cell its expression profile's closest match within the dataset linked from the paper linked [here](https://www.nature.com/articles/s41467-020-19012-4), via the use of SingleR

    -   Reference dataset is of neuroendocrine cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

#### Reference Dataset

```{r, eval = FALSE}

#Just run entire chunk

NeuroMeta <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_CellInfo.csv.gz")

NeuroExpression <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_5181_count.csv.gz")

columns_to_keep <- colnames(NeuroExpression) %in% NeuroMeta$X | colnames(NeuroExpression) == "X"

NeuroExpressionFiltered <- NeuroExpression[, columns_to_keep]

rm(NeuroExpression)

rownames(NeuroExpressionFiltered) <- NeuroExpressionFiltered[,1]
NeuroExpressionFiltered <- NeuroExpressionFiltered[,-1]

NeuroExpressionFiltered <- NeuroExpressionFiltered %>%
  mutate(across(everything(), as.numeric))

NeuroIDs <- as.character(NeuroMeta$cell_type)
names(NeuroIDs) <- colnames(NeuroExpressionFiltered)

rm(NeuroMeta)
NeuroExpressionFiltered <- as.matrix(NeuroExpressionFiltered)


```

#### Function

```{r, eval = FALSE}
SingleNeuro <- function(seuratObject) {
  # Function to run SingleR
  NeuroClusterLabels <- SingleR(test = as.SingleCellExperiment(seuratObject),
                                ref = NeuroExpressionFiltered,
                                labels = NeuroIDs)
  
  #To add SingleR output to Seurat Object
  seuratObject1 <- AddMetaData(seuratObject, metadata = NeuroClusterLabels@listData[["labels"]], col.name = "NeuroCellLabels")
  
  #To group everything by Neuro Cell Type
  Idents(seuratObject1) <- "NeuroCellLabels"
  return(seuratObject1)
}
```

#### Example:

```{r, eval = FALSE}
ex_sample10 <- SingleNeuro(ex_sample10)
```

```{r}
DimPlot(ex_sample10, 
        reduction = "umap",
        group.by = "NeuroCellLabels")
```

### SingleImmune

#### **Explanation:**

-   Annotates each cell identified based on its expression profiles' closest match within the dataset from the paper linked [here](https://www.science.org/doi/10.1126/science.abl5197), via the use of [SCINA](https://pmc.ncbi.nlm.nih.gov/articles/PMC6678337/).

-   Cell identities may be found via the dataframe `seuratObject@meta.data`, column = `ImmuneCellLabels`

-   **Note:** Notice that some of the cells from the previous SingleNeuro() function were labeled "Imm". These are typically the only cells that I ran this function on, so as to not mischaracterize previously identified neuroendocrine cells.

-   **Note:** SCINA is used in place of SingleR due to the different structure of the reference dataset.

-   **Note:** The function is named *`Single`*`Immune` purely for syntax consistency, as the function serves the same effective purpose as the aforementioned `SingleHPCA` despite using SCINA instead of SingleR.

    -   Reference dataset is of immune cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

#### Reference Dataset

```{r, eval = FALSE}
ImmuneRef <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Immune Cells/E-CURD-122.marker_genesymbols_inferred_cell_type_SCINA.csv")
ImmuneRef_list <- lapply(1:39, function(i) {
  genes <- ImmuneRef[, i]
  genes <- genes[!is.na(genes)]
  return(genes)
})

names(ImmuneRef_list) <- colnames(ImmuneRef)
ImmuneRef <- ImmuneRef_list
rm(ImmuneRef_list)
```

#### Function

```{r, eval = FALSE}
SingleImmune <- function(seuratObject) {
  
SeuratObjectExpressionData <- as.matrix(GetAssayData(seuratObject, assay = "RNA", layer = "data"))

#NOTE: significant overlap of genetic signatures between cell types--had to set rm_overlap to FALSE

results <- SCINA(SeuratObjectExpressionData, ImmuneRef, rm_overlap = FALSE)
seuratObject1 <- AddMetaData(seuratObject, metadata = results$cell_labels, col.name = "ImmuneCellLabels")
Idents(seuratObject1) <- "ImmuneCellLabels"
return(seuratObject1)
}
```

#### Example:

```{r, eval = FALSE}

#notice that we are leveraging our SingleNeuro characterizations
ex_sample10 <- SingleImmune(ex_sample10)
ex_sample10imm <- subset(ex_sample10, NeuroCellLabels == "Imm")

# this dataset results in a VERY large number of cell IDs, so for legibility, we are getting rid of any cell ID with less than 5 cells

CategoryCounts <- ex_sample10imm@meta.data %>%
  group_by(ImmuneCellLabels) %>%
  tally()

ValidCategories <- CategoryCounts %>% 
  filter(n >= 5) %>%
  pull(ImmuneCellLabels)

```

```{r}
DimPlot(subset(ex_sample10imm,  ImmuneCellLabels %in% ValidCategories), 
        reduction = "umap", 
        group.by = "ImmuneCellLabels")
```

### SingleUnbiasedNeuro

-   **Explanation:** Annotates each individual cluster previously identified via principal component analysis (aka via `FindCellClusters`) based on its collective expression profile's closest match within the dataset linked from the paper linked [here](https://www.nature.com/articles/s41467-020-19012-4), via the use of SingleR

    -   Reference dataset is of neuroendocrine cells (same as `SingleNeuro`)

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

    -   Annotations can be referenced via `seuratObject@meta.data$UnbiasedNeuroCellLabels`

```{r, eval = FALSE}
SingleUnbiasedNeuro <- function(seuratObject){
  
  # Create a table of `seurat_clusters` vs `NeuroCellLabels`
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$NeuroCellLabels)
  
  # Identify the most abundant label for each cluster
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] # Returns the label with the highest count
  })
  
  # Relabel `seurat_clusters` with the most abundant `NeuroCellLabels`
  seuratObject@meta.data$UnbiasedNeuroCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  return(seuratObject)
}
```

#### Example:

```{r, eval = FALSE}
ex_sample10 <- SingleUnbiasedNeuro(ex_sample10)
```

```{r}
# notice that it is each cluster within "seurat_clusters" that is relabeled 
DimPlot(ex_sample10, 
        reduction = "umap",
        group.by = "seurat_clusters")
DimPlot(ex_sample10, 
        reduction = "umap",
        group.by = "UnbiasedNeuroCellLabels")
```

### SingleUnbiasedImmune

-   **Explanation:** Annotates each individual cluster previously identified via principle component analysis (aka via `FindCellClusters`) based on its expression profiles' closest match within the dataset from the paper linked [here](https://www.science.org/doi/10.1126/science.abl5197), via the use of [SCINA](https://pmc.ncbi.nlm.nih.gov/articles/PMC6678337/).

-   **Note:** SCINA is used in place of SingleR due to the different structure of the reference dataset.

-   **Note:** The function is named *`Single`*`UnbiasedImmune` purely for syntax consistency, as the function serves the same effective purpose as the aforementioned `SingleUnbiasedNeuro`, despite using SCINA instead of SingleR.

    -   Reference dataset is of immune cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

```{r, eval = FALSE}
SingleUnbiasedImmune <- function(seuratObject){
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$ImmuneCellLabels)
  
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] 
  })
  
  seuratObject@meta.data$UnbiasedImmuneCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  
  return(seuratObject)
}
```

#### Example:

```{r, eval = FALSE}
# notice that we are using the Seurat object that is subsetted to only have immune cells (identified via SingleNeuro)
ex_sample10imm <- SingleUnbiasedImmune(ex_sample10imm)
```

```{r}
# notice that it is each cluster within "seurat_clusters" that is relabeled 
DimPlot(ex_sample10imm, 
        reduction = "umap",
        group.by = "seurat_clusters")
DimPlot(ex_sample10imm, 
        reduction = "umap",
        group.by = "UnbiasedImmuneCellLabels")

```

### BarPlotIdPercent

-   **Explanation:** Creates a barplot detailing the percentage each cell identity makes up of the total sample

    -   **Note:** the `Identity` parameter needs to be a character string in quotation marks for the function to work properly

-   **Parameters:**

    -   **Identity**: the cell annotation that will be used to make the barplot

        -   e.g. `"NeuroCellLabels"` or `"seurat_clusters"`

```{r, eval = FALSE}
#To create an inclusive barplot based on cell identities 
BarPlotIdPercent <- function(seuratObject, Identity){
  
  TotalCells <- dim(seuratObject@meta.data)[1]
  
  cellIdentities <- seuratObject@meta.data[[Identity]]
  
  cellCounts <- table(cellIdentities)
  
  cellCountsDF <- as.data.frame(cellCounts)
  
  colnames(cellCountsDF) <- c("CellIdentity", "Count")
  
  ggplot(cellCountsDF, aes(x = CellIdentity, 
                           y = Count/TotalCells*100, 
                           fill = CellIdentity)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + 
    labs(title = "Distribution of Cell Types", x = "Cell Identity", y = "Percentage of Cells") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme(text = element_text("sans"))
}

```

#### Example:

```{r}
BarPlotIdPercent(ex_sample10, "NeuroCellLabels")
```
