---
title: "Erson Analysis"
output: html_document
date: "2025-02-06"
author: "Sebastian Reyes"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

The below framework outlines a workflow I designed to streamline the steps outlined in this [Seurat vignette](https://satijalab.org/seurat/articles/pbmc3k_tutorial) for the purposes needed by the Erson Lab in the summer of 2024--when I made this. Really, the entire framework below is a reframing of this vignette to better suit the specific needs of the Erson Lab—from the perspective of my specific project, anyway.

To explain the "custom functions" created below: many of these functions are used repeatedly with different parameters during the exploratory analyses I conducted on the Erson Lab's samples. Making these custom functions quickened the workflow and allowed parameters to more easily be edited.

That said, reading the aforementioned [Seurat vignette](https://satijalab.org/seurat/articles/pbmc3k_tutorial) would provide the reader with the vast majority of the information contained in the below outline.

## 2. Installing Packages Needed

```{r, eval = FALSE}
install.packages("dplyr")
install.packages("Seurat")
install.packages("patchwork")
install.packages("tibble")
install.packages("SingleR")
install.packages("SCINA")
BiocManager::install("celldex")
install.packages("ggplot2")
```

## 3. Loading Packages Needed

```{r, eval = FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(tibble)
library(SingleR)
library(SCINA)
library(celldex)
library(ggplot2)
```

## 4. All Custom Functions Used

### CreateSeuratObject

-   **Explanation**: more easily creates a Seurat object based on file directory

-   **Parameters**:

    -   fileDirectory: file path (must be h5 file for this specific function)

    -   objectName: self-explanatory

    -   projectName: self-explanatory

```{r, eval = FALSE}
CreateSeuratObject <- function(fileDirectory, objectName, projectName) {
  rawCounts <- Read10X_h5(fileDirectory)
  seurat_obj <- Seurat::CreateSeuratObject(counts = rawCounts, project = projectName, min.cells = 3, min.features = 200)
  assign(objectName, seurat_obj, envir = .GlobalEnv)
  return(objectName)
}
```

### MakeVlnPlot

-   **Explanation**: more easily creates violin plots for QC purposes

    -   nFeature_RNA = gives the number of unique features (genes) in a given cell, each point represents one cell

    -   nCount_RNA = gives the total number of RNA molecules in a cell, each point represents one cell

    -   percent.mt = percentage of mitochondrial RNA, each point represents one cell

```{r, eval = FALSE}
MakeVlnPlot <- function(seuratObject) {
  print(VlnPlot(seuratObject, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3))
  return(seuratObject)
}
```

### AddMitoRna

-   **Explanation**: appends mitochondrial RNA data to Seurat object on a cell-by-cell basis

```{r, eval = FALSE}
AddMitoRna <- function(seuratObject){
  seuratObject[["percent.mt"]] <- Seurat::PercentageFeatureSet(seuratObject, pattern = "^MT-")
  return(seuratObject)
}
```

### QCSeurat

-   **Explanation:** streamlines the quality control process, allows for low-quality cells to be eliminated based on feature count and mitochondrial RNA percentage

-   **Parameters:**

    -   seuratObject = seuratObject to be trimmed

    -   minFeature = minimum number of genes a cell must contain to remain included in data

    -   maxFeature = maximum number of genes a cell can contain to remain included in data

    -   percentMT = maximum percentage of mitochondrial RNA which a cell can contain to remain included in data

```{r, eval = FALSE}
QCSeurat <- function(seuratObject, minFeature, maxFeature, percentMT) {
  seuratObject <- subset(seuratObject, subset = nFeature_RNA > minFeature & nFeature_RNA < maxFeature & percent.mt < percentMT)
  return(seuratObject)
}
```

### MakeFeaturePlot

-   **Explanation**: Makes scatterplot showing correlation between 2 features (genes) of a seurat object

```{r, eval = FALSE}
MakeFeaturePlot <- function(seuratObject, feature1, feature2) {
  plot1 <- FeatureScatter(seuratObject, feature1 = feature1, feature2 = feature2)
  print(plot1)
}
```

### NormalizeAndVariable

-   **Explanation**: normalizes data, finds genes within Seurat object that display variable expression patterns, and creates a plot highlighting the top 10 genes with the highest degree of variability in their expression.

```{r, eval = FALSE}
NormalizeAndVariable <- function(seuratObject) {
  seuratObject <- NormalizeData(seuratObject)
  seuratObject <- FindVariableFeatures(seuratObject, selection.method = "vst", nfeatures = 2000)
  top10 <- head(VariableFeatures(seuratObject), 5)
  plot1 <- VariableFeaturePlot(seuratObject)
  plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
  print(plot2)
  return(seuratObject)
}
```

### Scale

-   **Explanation**: scales the Seurat object as per the workflow outlined in the Seurat vignette

```{r, eval = FALSE}
Scale <- function(seuratObject) {
  seuratObject <- ScaleData(seuratObject, features = rownames(seuratObject))
  seuratObject <- ScaleData(seuratObject, vars.to.regress = "percent.mt",)
  return(seuratObject)
}
```

### PcaElbow

-   **Explanation:** Runs a principal component analysis and creates an elbow plot based on it

```{r, eval = FALSE}
PcaElbow <- function(seuratObject) {
  seuratObject <- RunPCA(seuratObject, features = VariableFeatures(seuratObject))
  print(ElbowPlot(object = seuratObject))
  return(seuratObject)
}
```

### MakeHeatmap

-   **Explanation:** Creates one heatmap for every principal component included in `PcNumberVector`

    -   `PcNumberVector` should be a numeric vector

```{r, eval = FALSE}
MakeHeatmap <- function(seuratObject, PcNumberVector){
  DimHeatmap(seuratObject, dims = PcNumberVector, balanced = TRUE)
  return(seuratObject)
}
```

### FindCellClusters

-   **Explanation:** finds clusters of cells based of principal component behavior

-   **Parameters**:

    -   seuratObject = self-explanatory

    -   PCs = a numeric vector containing the principal components to be used

    -   resolution = a value between 0 and 1; defines the granularity of the clusters

-   Cluster assignments may be found via the dataframe `seuratObject@meta.data`, column = `seurat_clusters`

```{r, eval = FALSE}
FindCellClusters <- function(seuratObject, PCs, resolution) {
  seuratObject2 <- FindNeighbors(seuratObject, dims = PCs)
  seuratObject3 <- FindClusters(seuratObject2, resolution = resolution)
  return(seuratObject3)
}
```

### MakeUMAP

-   **Explanation:** makes a UMAP plot based on the specified principal components

-   **Parameters:**

    -   seuratObject = self-explanatory

    -   PCs = a numeric vector containing the principal components to be used

```{r, eval = FALSE}
MakeUMAP <- function(seuratObject1, PCs) {
  seuratObject2 <- RunUMAP(seuratObject1, dims = PCs)
  print(DimPlot(seuratObject2, reduction = "umap"))
  return(seuratObject2)
}
```

### FindPosMarkers

-   **Explanation:** Finds markers that are upregulated within previously established clusters

```{r, eval = FALSE}
FindPosMarkers <- function(seuratObject){
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  posMarkers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC >1)
  return(seuratObject)
}
```

### Top5Markers

-   **Explanation:** Prints out the top 5 most upregulated genetic markers, on a cluster-by-cluster basis

```{r, eval = FALSE}
Top5Markers <- function(seuratObject) {
  posMarkers <- FindAllMarkers(seuratObject, only.pos = TRUE)
  separatedMarkers <- posMarkers %>%
    dplyr::filter(avg_log2FC >1) %>%
    group_by(cluster) %>%
    slice_head(n = 5)
  print(separatedMarkers, n = Inf)
}
```

### SingleHPCA

-   **Explanation:** Annotates each individual cell based on its expression profile's closest match within the Human Primary Cell Atlas reference dataset; based on the use of SingleR

    -   HPCA dataset is more generic than the below neuroendocrine or immune cell datasets, but is good for establishing an initial tissue composition overview

    -   Cell identities may be found via the dataframe `seuratObject@meta.data`, column = `HPCACellLabels`

```{r, eval = FALSE}
SingleHPCA <- function(seuratObject){
  HpcaLabels <- SingleR(test = as.SingleCellExperiment(seuratObject), 
                        ref =    HumanPrimaryCellAtlasData(), 
                        labels = HumanPrimaryCellAtlasData()$label.main)
  seuratObject1 <- AddMetaData(seuratObject, 
                               metadata = HpcaLabels@listData[["labels"]], 
                               col.name = "HPCACellLabels")
  return(seuratObject1)
}
```

### SingleImmune

-   **Explanation:** Annotates each cell identified based on its expression profiles' closest match within the dataset from the paper linked [here](https://www.science.org/doi/10.1126/science.abl5197), via the use of [SCINA](https://pmc.ncbi.nlm.nih.gov/articles/PMC6678337/).

-   **Note:** SCINA is used in place of SingleR due to the different structure of the reference dataset.

-   **Note:** The function is named *`Single`*`Immune` purely for syntax consistency, as the function serves the same effective purpose as the aforementioned `SingleHPCA` despite using SCINA instead of SingleR.

    -   Reference dataset is of immune cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

#### Reference Dataset

```{r, eval = FALSE}
ImmuneRef <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Immune Cells/E-CURD-122.marker_genesymbols_inferred_cell_type_SCINA.csv")
ImmuneRef_list <- lapply(1:39, function(i) {
  genes <- ImmuneRef[, i]
  genes <- genes[!is.na(genes)]
  return(genes)
})

names(ImmuneRef_list) <- colnames(ImmuneRef)
ImmuneRef <- ImmuneRef_list
rm(ImmuneRef_list)
```

#### Function

```{r, eval = FALSE}
SingleImmune <- function(seuratObject) {
  
SeuratObjectExpressionData <- as.matrix(GetAssayData(seuratObject, assay = "RNA", layer = "data"))

#NOTE: significant overlap of genetic signatures between cell types--had to set rm_overlap to FALSE

results <- SCINA(SeuratObjectExpressionData, ImmuneRef, rm_overlap = FALSE)
seuratObject1 <- AddMetaData(seuratObject, metadata = results$cell_labels, col.name = "ImmuneCellLabels")
Idents(seuratObject1) <- "ImmuneCellLabels"
return(seuratObject1)
}
```

### SingleNeuro

-   **Explanation:** Annotates each individual cell its expression profile's closest match within the dataset linked from the paper linked [here](https://www.nature.com/articles/s41467-020-19012-4), via the use of SingleR

    -   Reference dataset is of neuroendocrine cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

#### Reference Dataset

```{r, eval = FALSE}

#Just run entire chunk

NeuroMeta <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_CellInfo.csv.gz")

NeuroExpression <- read.csv("~/Documents/Erson_Lab/Datasets/CellIDReferences/Neuroendocrine Cells/GSE142653_pit_dev_5181_count.csv.gz")

columns_to_keep <- colnames(NeuroExpression) %in% NeuroMeta$X | colnames(NeuroExpression) == "X"

NeuroExpressionFiltered <- NeuroExpression[, columns_to_keep]

rm(NeuroExpression)

rownames(NeuroExpressionFiltered) <- NeuroExpressionFiltered[,1]
NeuroExpressionFiltered <- NeuroExpressionFiltered[,-1]

NeuroExpressionFiltered <- NeuroExpressionFiltered %>%
  mutate(across(everything(), as.numeric))

NeuroIDs <- as.character(NeuroMeta$cell_type)
names(NeuroIDs) <- colnames(NeuroExpressionFiltered)

rm(NeuroMeta)
NeuroExpressionFiltered <- as.matrix(NeuroExpressionFiltered)


```

#### Function

```{r, eval = FALSE}
SingleNeuro <- function(seuratObject) {
  # Function to run SingleR
  NeuroClusterLabels <- SingleR(test = as.SingleCellExperiment(seuratObject),
                                ref = NeuroExpressionFiltered,
                                labels = NeuroIDs)
  
  #To add SingleR output to Seurat Object
  seuratObject1 <- AddMetaData(seuratObject, metadata = NeuroClusterLabels@listData[["labels"]], col.name = "NeuroCellLabels")
  
  #To group everything by Neuro Cell Type
  Idents(seuratObject1) <- "NeuroCellLabels"
  return(seuratObject1)
}
```

### SingleUnbiasedNeuro

-   **Explanation:** Annotates each individual cluster previously identified via principal component analysis (aka via `FindCellClusters`) based on its expression profiles' closest match within the dataset linked from the paper linked [here](https://www.nature.com/articles/s41467-020-19012-4), via the use of SingleR

    -   Reference dataset is of neuroendocrine cells (same as `SingleNeuro`)

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

```{r, eval = FALSE}
SingleUnbiasedNeuro <- function(seuratObject){
  
  # Create a table of `seurat_clusters` vs `NeuroCellLabels`
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$NeuroCellLabels)
  
  # Identify the most abundant label for each cluster
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] # Returns the label with the highest count
  })
  
  # Relabel `seurat_clusters` with the most abundant `NeuroCellLabels`
  seuratObject@meta.data$UnbiasedNeuroCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  return(seuratObject)
}
```

### SingleUnbiasedImmune

-   **Explanation:** Annotates each individual cluster previously identified via principle component analysis (aka via `FindCellClusters`) based on its expression profiles' closest match within the dataset from the paper linked [here](https://www.science.org/doi/10.1126/science.abl5197), via the use of [SCINA](https://pmc.ncbi.nlm.nih.gov/articles/PMC6678337/).

-   **Note:** SCINA is used in place of SingleR due to the different structure of the reference dataset.

-   **Note:** The function is named *`Single`*`UnbiasedImmune` purely for syntax consistency, as the function serves the same effective purpose as the aforementioned `SingleUnbiasedNeuro`, despite using SCINA instead of SingleR.

    -   Reference dataset is of immune cells

    -   Reference dataset file is in the GitHub repository and should automatically be utilized when calling this function

```{r, eval = FALSE}
SingleUnbiasedImmune <- function(seuratObject){
  # Create a contingency table of `seurat_clusters` vs `NeuroCellLabels`
  cluster_label_counts <- table(seuratObject@meta.data$seurat_clusters, 
                                seuratObject@meta.data$ImmuneCellLabels)
  
  # Identify the most abundant label for each cluster
  most_abundant_labels <- apply(cluster_label_counts, 1, function(x) {
    names(x)[which.max(x)] # Returns the label with the highest count
  })
  
  # Relabel seurat_clusters with the most abundant NeuroCellLabel within them
  seuratObject@meta.data$UnbiasedImmuneCellLabels <- factor(seuratObject@meta.data$seurat_clusters, levels = names(most_abundant_labels), labels = most_abundant_labels)
  
  return(seuratObject)
}
```

### BarPlotIdPercent

-   **Explanation:** Creates a barplot detailing the percentage each cell identity makes up of the total sample

    -   **Note:** the `Identity` parameter needs to be a character string in quotation marks for the function to work properly

-   **Parameters:**

    -   **Identity**: the cell annotation that will be used to make the barplot

        -   e.g. `"NeuroCellLabels"` or `"seurat_clusters"`

```{r, eval = FALSE}
#To create an inclusive barplot based on cell identities 

BarPlotIdPercent <- function(seuratObject, Identity){
  
  TotalCells <- dim(seuratObject@meta.data)[1]
  
  cellIdentities <- seuratObject@meta.data[[Identity]]
  
  cellCounts <- table(cellIdentities)
  
  cellCountsDF <- as.data.frame(cellCounts)
  
  colnames(cellCountsDF) <- c("CellIdentity", "Count")
  
  ggplot(cellCountsDF, aes(x = CellIdentity, 
                           y = Count/TotalCells*100, 
                           fill = CellIdentity)) + 
    geom_bar(stat = "identity") +
    theme_minimal() + 
    labs(title = "Distribution of Cell Types", x = "Cell Identity", y = "Percentage of Cells") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    theme(text = element_text("sans"))
}

```

## 5. Samples

### Sample 1

#### Initial QC

```{r, eval = FALSE}
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA1/DATA1.H5",
                   "sample1", 
                   "sample1")
sample1 <- AddMitoRna(sample1)
MakeVlnPlot(sample1)
sample1 <- QCSeurat(sample1, 200, 3000, 25)
MakeFeaturePlot(sample1, "nCount_RNA", "percent.mt")
MakeFeaturePlot(sample1, "nCount_RNA", "nFeature_RNA")
sample1 <- NormalizeAndVariable(sample1)
sample1 <- Scale(sample1)
sample1 <- PcaElbow(sample1)
MakeHeatmap(sample1, 1:8)
```

#### General Exploratory Analyses

```{r, eval = FALSE}
sample1 <- FindCellClusters(sample1, 1:7, 0.5)
sample1 <- RunUMAP(sample1, dims = 1:7)
sample1 <- FindPosMarkers(sample1)
sample1 <- SingleHPCA(sample1)
sample1 <- SingleNeuro()
sample1 <- SingleImmune(sample1)
```

#### Specific Exploratory Analysis/Visualizations

```{r, eval = FALSE}
#For immune characterization
    SeuratFiltered <- subset(sample1, subset = NeuroCellLabels == "Imm")
    CategoryCounts <- SeuratFiltered@meta.data %>%
      group_by(ImmuneCellLabels) %>%
      tally()
    ValidCategories <- CategoryCounts %>% 
      filter(n >= 5) %>%
      pull(ImmuneCellLabels)
    
    MakeUMAP(sample1, 1:7)
    
print(DimPlot(subset(sample1, subset = NeuroCellLabels == "Imm" & ImmuneCellLabels %in% ValidCategories, cols = MyColors), reduction = "umap", group.by = "ImmuneCellLabels", label = FALSE) + ggtitle("Immune Cell Types"))
Idents(sample1) <- sample1@meta.data$NeuroCellLabels
BarPlot(sample1, 8224) + ggtitle("Unbiased Neuroendocrine Cell Type Distribution")

```

### Sample 2

```{r, eval = FALSE}
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA2/filtered_feature_bc_matrix.h5",
                   "sample2", 
                   "sample2" )
sample2 <- AddMitoRna(sample2)
MakeVlnPlot(sample2)
dim(sample2)
sample2 <- QCSeurat(sample2, 200, 5000, 30)
dim(sample2)
MakeFeaturePlot(sample2, "nCount_RNA", "nFeature_RNA")
sample2 <- NormalizeAndVariable(sample2)
sample2 <- Scale(sample2)
sample2 <- PcaElbow(sample2)
sample2 <- RunUMAP(sample2, dims = 1:9)
sample2 <- FindCellClusters(sample2, 1:9, 0.5)
DimPlot(sample2, reduction = "umap", label = TRUE) 
sample2 <- FindPosMarkers(sample2)
sample2 <- SingleUnbiasedNeuro(sample2)

    #For immune characterization
    SeuratFiltered <- subset(sample2, subset = NeuroCellLabels == "Imm")
    CategoryCounts <- SeuratFiltered@meta.data %>%
      group_by(ImmuneCellLabels) %>%
      tally()
    ValidCategories <- CategoryCounts %>% 
      filter(n >= 5) %>%
      pull(ImmuneCellLabels)

print(DimPlot(subset(sample2, subset = NeuroCellLabels == "Imm" & ImmuneCellLabels %in% ValidCategories), reduction = "umap", group.by = "ImmuneCellLabels") + ggtitle("Sample 2 Immune Cell Types"))
Idents(sample2) <- sample2@meta.data$UnbiasedNeuroCellLabels
BarPlot(sample2, 7359) + ggtitle("Unbiased Neuroendocrine Cell Type Distribution")
```

### Sample 3

```{r, eval = FALSE}
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA3/filtered_feature_bc_matrix.h5",
                   "sample3", 
                   "sample3")
sample3 <- AddMitoRna(sample3)
MakeVlnPlot(sample3)
dim(sample3)
sample3 <- QCSeurat(sample3, 200, 4000, 40)
dim(sample3)
MakeFeaturePlot(sample3, "nCount_RNA", "percent.mt")
MakeFeaturePlot(sample3, "nCount_RNA", "nFeature_RNA")
sample3 <- NormalizeAndVariable(sample3)
sample3 <- Scale(sample3)
sample3 <- PcaElbow(sample3)
MakeHeatmap(sample3, 1:10)
sample3 <- FindCellClusters(sample3, 1:10, 1)
sample3 <- RunUMAP(sample3, dims = 1:10)
DimPlot(sample3, reduction = "umap", label = TRUE, group.by = "UnbiasedNeuroCellLabels")
sample3 <- FindPosMarkers(sample3)
sample3 <- SingleUnbiasedNeuro(sample3)

#For immune characterization
SeuratFiltered <- subset(sample3, subset = NeuroCellLabels == "Imm")
CategoryCounts <- SeuratFiltered@meta.data %>%
  group_by(ImmuneCellLabels) %>%
  tally()
ValidCategories <- CategoryCounts %>% 
  filter(n >= 5) %>%
  pull(ImmuneCellLabels)

print(DimPlot(subset(sample3, subset = NeuroCellLabels == "Imm" & ImmuneCellLabels %in% ValidCategories), reduction = "umap", group.by = "ImmuneCellLabels") + ggtitle("Sample 3 Immune Cell Types"))
Idents(sample3) <- sample3@meta.data$UnbiasedNeuroCellLabels
BarPlot(sample3, 6200) + ggtitle("Unbiased Neuroendocrine Cell Type Distribution")

```

### Sample 4

```{r, eval = FALSE}
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA4/filtered_feature_bc_matrix.h5",
                   "sample4", 
                   "sample4" )
sample4 <- AddMitoRna(sample4)
MakeVlnPlot(sample4)
dim(sample4)
sample4 <- QCSeurat(sample4, 200, 5000, 7)
dim(sample4)
MakeFeaturePlot(sample4, "nCount_RNA", "percent.mt")
MakeFeaturePlot(sample4, "nCount_RNA", "nFeature_RNA")
sample4 <- NormalizeAndVariable(sample4)
sample4 <- Scale(sample4)
sample4 <- PcaElbow(sample4)
MakeHeatmap(sample4, 1:11)
sample4 <- FindCellClusters(sample4, 1:11, 1)
sample4 <- FindPosMarkers(sample4)
sample4 <- SingleUnbiasedNeuro(sample4)
sample4 <- RunUMAP(sample4, dims = 1:11)
print(DimPlot(sample4, reduction = "umap", group.by = "UnbiasedNeuroCellLabels", label = TRUE) + ggtitle("Unbiased Neuroendocrine Cell Types"))
Idents(sample4) <- sample4@meta.data$UnbiasedNeuroCellLabels
    
    #for immune characterization
    SeuratFiltered <- subset(sample4, subset = NeuroCellLabels == "Imm")
    CategoryCounts <- SeuratFiltered@meta.data %>%
      group_by(ImmuneCellLabels) %>%
      tally()
    ValidCategories <- CategoryCounts %>% 
      filter(n >= 1) %>%
      pull(ImmuneCellLabels)

BarPlot(sample4, 14986) + ggtitle("Unbiased Neuroendocrine Cell Type Distribution")
BarPlot(sample4, 14986)
VlnPlot(sample4, features = "STAT6") +  NoLegend() + ggtitle("STAT6 (P3/CD/Tumor)")
```

### Sample 5

```{r, eval = FALSE}
#Sample 5
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA5/filtered_feature_bc_matrix.h5",
                   "sample5", 
                   "sample5" )
sample5 <- AddMitoRna(sample5)
MakeVlnPlot(sample5)
dim(sample5)
sample5 <- QCSeurat(sample5, 200, 2000, 40)
dim(sample5)
MakeFeaturePlot(sample5, "nCount_RNA", "percent.mt")
MakeFeaturePlot(sample5, "nCount_RNA", "nFeature_RNA")
sample5 <- NormalizeAndVariable(sample5)
sample5 <- Scale(sample5)
sample5 <- PcaElbow(sample5)
MakeHeatmap(sample5, 1:12)
sample5 <- FindCellClusters(sample5, 1:9, 1.5)
sample5 <- FindPosMarkers(sample5)
sample5 <- SingleUnbiasedNeuro(sample5)
sample5 <- RunUMAP(sample5, dims = 1:9)
print(DimPlot(sample5, reduction = "umap", group.by = "UnbiasedNeuroCellLabels") + ggtitle("Unbiased Neuroendocrine Cell Types"))
Idents(sample5) <- sample5@meta.data$UnbiasedNeuroCellLabels

#for immune characterization
SeuratFiltered <- subset(sample5, subset = NeuroCellLabels == "Imm")
CategoryCounts <- SeuratFiltered@meta.data %>%
  group_by(ImmuneCellLabels) %>%
  tally()
ValidCategories <- CategoryCounts %>% 
  filter(n >= 1) %>%
  pull(ImmuneCellLabels)
BarPlot(sample5, 1638) + ggtitle("Neuroendocrine Cell Distribution")
```

### Sample 6 (Omitted)

### Sample 7

```{r, eval = FALSE}
CreateSeuratObject("~/Documents/Erson_Lab/Datasets/Erson Samples/PA_SC_SR/DATA7/filtered_feature_bc_matrix.h5",
                   "sample7", 
                   "sample7")
sample7 <- AddMitoRna(sample7)
MakeVlnPlot(sample7)
dim(sample7)
sample7 <- QCSeurat(sample7, 200, 2000, 9)
dim(sample7)
MakeFeaturePlot(sample7, "nCount_RNA", "percent.mt")
MakeFeaturePlot(sample7, "nCount_RNA", "nFeature_RNA")
sample7 <- NormalizeAndVariable(sample7)
sample7 <- Scale(sample7)
sample7 <- PcaElbow(sample7)
MakeHeatmap(sample7, 1:12)
sample7 <- FindCellClusters(sample7, 1:9, 1)
sample7 <- FindPosMarkers(sample7) 
sample7 <- SingleUnbiasedNeuro(sample7)
sample7 <- RunUMAP(sample7, dims = 1:9)
print(DimPlot(sample7, reduction = "umap", group.by = "UnbiasedNeuroCellLabels", label = TRUE) + ggtitle("Unbiased Neuroendocrine Cell Types"))
Idents(sample7) <- sample7@meta.data$UnbiasedNeuroCellLabels
BarPlot(sample7, 10370) + ggtitle("Unbiased Neuroendocrine Cell Types")

#for immune characterization
SeuratFiltered <- subset(sample7, subset = NeuroCellLabels == "Imm")
CategoryCounts <- SeuratFiltered@meta.data %>%
    group_by(ImmuneCellLabels) %>%
    tally()
    ValidCategories <- CategoryCounts %>% 
    filter(n >= 1) %>%
    pull(ImmuneCellLabels)
BarPlot(subset(sample7, subset = NeuroCellLabels == "Imm"), 10370) + ggtitle("Immune Cell Distribution")
```
